Irreducible Components and Dimension of the Springer Fiber of a Hook-Type Slodowy Slice

### 2024-09-01 paper.tex
read through section 2 up to theorem 2.2

\section{Preliminary Definitions and Facts}\label{prelim}
\subsection{Conventions and Notations}
We write \(\GL_m, \spl_m\) to denote \(\GL_m(\mathbb{C}), \spl_m(\mathbb{C})\), and so on.
By \(J_m\) we refer to the nilpotent \(m \times m\) Jordan block (which, by convention, has ones \emph{below} the diagonal).
Given a partition \(\lambda = (\lambda_1, ..., \lambda_k)\), we write \(J(\lambda)\) to denote the block matrix
\[\begin{pmatrix}
    J_{\lambda_1} \\
    & \ddots \\
    & & J_{\lambda_k}
\end{pmatrix}.\]
A partition is always indexed in nonincreasing order, even if it is defined differently.
For example, if \(\mu = (4, 6, 3)\), then \(\mu_1 = 6, \mu_2 = 4, \mu_3 = 3\).

\subsection{Springer fibers}
Let \(\mathfrak{g}\) be a Lie algebra.
Let \(\mathcal{N} \subseteq \mathfrak{g}\) be the subset consisting of nilpotent elements.
Let \(\mathcal{B}\) be the variety of Borel subalgebras of \(\mathfrak{g}\).
Let \(\widetilde{\mathcal{N}} = \{(\mathfrak{b}, n) : n \in \mathfrak{b}\} \subseteq \mathcal{B} \times \mathcal{N}\).
Let \(\pi : \widetilde{\mathcal{N}} \to \mathcal{N}\) be the projection onto the second coordinate.
We call this the \emph{Springer resolution}.
For \(n \in \mathcal{N}\), we call \(\pi^{-1}(n)\) the \emph{Springer fiber at \(n\)}.

\subsection{Springer fibers in \texorpdfstring{\(\gl_m\)}{gl\_m}}
Now we let \(\mathcal{N}_m\) be the set of nilpotent elements in \(\gl_m\), and \(\mathcal{B}_m\) the variety of Borel subalgebras of \(\gl_m\).
Let \(H \subseteq \gl_m\) be the subalgebra of upper triangular matrices.
The variety of Borel subalgebras of \(\gl_m\) is \(\mathcal{B} = \{gHg^{-1} : g \in \GL_m\}\).
Thus, the Springer fiber at \(X \in \mathcal{N}\) is 
\[\mathcal{F}_X = \{gHg^{-1} : X \in gHg^{-1}\}.\]
% ^ use that notation everywhere for springer fiber
\begin{definition}
    A \emph{flag} \((V_i)\) of \(\mathbb{C}^m\) is a sequence of subspaces
    \[0 = V_0 \subseteq V_1 \subseteq \cdots \subseteq V_m = \mathbb{C}^m,\]
    where \(\dim V_i = i\).
\end{definition}
We say that \(X \in \gl_m\) \emph{preserves} a flag \((V_i)\) if \(\forall i. \; XV_i \subseteq V_i\).
Note that \(X \in \mathcal{N}\) preserves a flag \((V_i)\) if and only if \(\forall i. \; XV_i \subseteq V_{i - 1}\).
\par The simplest flag is the \emph{standard flag} \((E_i)\), where \(E_i := \langle e_1, ..., e_i \rangle\).
Note that the group \(H\) is exactly the subset of \(\gl_m\) which preserves \(E_i\).
\par We can think of \(\mathcal{B}\) as the set of flags of \(\mathbb{C}^m\), via the correspondence 
\[gHg^{-1} \leftrightarrow (gE_i).\]
Note that \(X\) preserves \((gE_i)\) if and only if \(X \in gHg^{-1}\).
Thus, we may write the Springer fiber at \(X \in \mathcal{N}\) in terms of flags, as 
\[S_X = \{(gE_i) : X \in gHg^{-1}\} = \{(V_i) : \forall i. \; XV_i \subseteq V_{i - 1}\}.\]

>>>>

\section{Preliminary Definitions and Facts}\label{prelim}
\subsection{Conventions and Notations}
We write \(\GL_m, \spl_m\) to denote \(\GL_m(\mathbb{C}), \spl_m(\mathbb{C})\), and so on.
By \(J_m\) we refer to the nilpotent \(m \times m\) Jordan block (which has ones \emph{below} the diagonal).
Given a partition \(\lambda = (\lambda_1, ..., \lambda_\ell)\), we write \(J(\lambda)\) to denote the block matrix
\[\begin{pmatrix}
    J_{\lambda_1} \\
    & \ddots \\
    & & J_{\lambda_\ell}
\end{pmatrix}.\]
A partition is always indexed in nonincreasing order, even if it is written down differently when defined.
For example, if \(\mu = (4, 6, 3)\), then \((\mu_1, \mu_2, \mu_3) = (6, 4, 3)\).
\par Throughout the paper, when a nilpotent operator \(X : \mathbb{C}^m \to \mathbb{C}^m\) of shape \(\lambda = (\lambda_1, ..., \lambda_\ell) \vdash m\) is in the context, we write \((e_{ij})_{1 \leq i \leq \ell, 1 \leq j \leq \lambda_i}\) to denote a Jordan basis of \(\mathbb{C}^m\) with the property that for all \(i,j\) we have \(Xe_{ij} = e_{i,j - 1}\).
(By convention, \(e_{ij} := 0\) for \(j < 1\), so that condition makes sense.)
Also, we will write \(f_n, ..., f_1\) to denote the basis of \(\mathbb{C}^n\) with the property that \(J_n f_i = f_{i - 1}\).

\subsection{Springer fibers}
Let \(\mathfrak{h}\) be a Lie algebra, let \(\mathcal{N}_\mathfrak{h} \subseteq \mathfrak{h}\) be the nilpotent cone in \(\mathfrak{h}\), and let \(\mathcal{B}_\mathfrak{h}\) be the variety of Borel subalgebras of \(\mathfrak{h}\).
Let \(\widetilde{\mathcal{N}}_\mathfrak{h} = \{(\mathfrak{b}, n) \in \mathcal{B}_\mathfrak{h} \times \mathcal{N}_\mathfrak{h} : n \in \mathfrak{b}\}\).
Let \(\pi_\mathfrak{h} : \widetilde{\mathcal{N}}_\mathfrak{h} \to \mathcal{N}_\mathfrak{h}\) be the projection onto the second coordinate.
We call this the \emph{Springer resolution}.
For \(n \in \mathcal{N}_\mathfrak{h}\), we call \(\pi_\mathfrak{h}^{-1}(n)\) the \emph{Springer fiber at \(n\)}.

\subsection{Springer fibers in \texorpdfstring{\(\gl_m\)}{gl\_m}}
Let \(\mathcal{N}_m\) be the nilpotent cone in \(\gl_m\), and let \(\mathcal{B}_m\) the variety of Borel subalgebras of \(\gl_m\).
Let \(\mathfrak{h} \subseteq \gl_m\) be the subalgebra of upper triangular matrices.
The variety of Borel subalgebras of \(\gl_m\) is \(\mathcal{B}_m = \{g\mathfrak{h}g^{-1} : g \in \GL_m\}\).
Thus, the Springer fiber at \(X \in \mathcal{N}_m\) is 
\[\mathcal{F}_X = \{g\mathfrak{h}g^{-1} : g \in \GL_m; X \in g\mathfrak{h}g^{-1}\}.\]
\begin{definition}
    A \emph{flag} \(V_\bullet\) of \(\mathbb{C}^m\) is a sequence of subspaces
    \[0 = V_0 \subseteq V_1 \subseteq \cdots \subseteq V_m = \mathbb{C}^m,\]
    where \(\dim V_i = i\).
\end{definition}
We say that \(X \in \gl_m\) \emph{preserves} a flag \(V_\bullet\) if for each \(i\) we have \(XV_i \subseteq V_i\).
Note that for \(X \in \gl_m\), we have [\(X \in \mathcal{N}_m\), and \(X\) preserves \(V_\bullet\)] if and only if [for all \(i\), \(XV_{i + 1} \subseteq V_i\)].
\par One flag is the \emph{standard flag} \(E_\bullet\), where \(E_i\) is generated by the first \(i\) standard basis vectors.
The subalgebra \(\mathfrak{h}\) is exactly the subset of \(\gl_m\) which preserves \(E_\bullet\).
So, we can think of \(\mathcal{B}_m\) as the variety of flags of \(\mathbb{C}^m\), via the correspondence 
\[g\mathfrak{h}g^{-1} \leftrightarrow g \cdot E_\bullet.\]
(\(\GL_n\) acts on the set of flags via \((g \cdot V_\bullet)_j = gV_j\).)
Note that \(X\) preserves \(g \cdot E_\bullet\) if and only if \(X \in g \mathfrak{h} g^{-1}\).
Thus, we may write the Springer fiber at \(X\) in terms of flags, as 
\[\mathcal{F}_X = \{g \cdot E_\bullet : X \in gHg^{-1}\} = \{V_\bullet : \forall i. \; XV_i \subseteq V_{i - 1}\}.\]
